What We Eval in the Shadows

### 2021-08-12 paper/main.tex
PR 5.3

Where does the expression passed to \eval come from? There are various means of
creating an expression, often these means correlate with a particular use case.
Our analysis records the values returned by some functions of interest as well
as their arguments. We build a direct acyclic graph that track the provenance of
the values that are given to \eval. The following example yields the provenance
graph with five nodes on Figure~\ref{fig:provgraph} for expression \k{e}.

\begin{lstlisting}
 > e <- parse(text = "a;b")
 > e[[2]] <- quote(c)
 > eval(e)
\end{lstlisting}

%\begin{figure}[h]
%\centering
%\includegraphics[width=8cm]{provenance-graph}
%\caption{The provenance DAG for the code example above. Assignments and complex assignments are intermediate nodes, whereas expression builders are leaves of the graph.} \label{fig:provgraph}
%\end{figure}

\begin{figure}[h]
	\centering
	\begin{tikzpicture}[
		every node/.style={draw, rounded corners, fill=LightGray, text width = 2cm, align = center, thick}
		]
		\node (E) at (2,0) {\k{eval(e)}};
		\node[text width = 3cm] (CA) at (2, 1.5) {\k{e[[2]] <-quote(c)}};
		\node (Q) at (4, 3) {\k{quote(c)}};
		\node[text width = 4cm] (A) at (0, 3) {\k{e <- parse(text = "a;b")}};
		\node[text width = 3cm] (P) at (0, 4.5) {\k{parse(text = "a;b")}};

		\draw[<-, >=latex] (E) -- (CA);
		\draw[<-, >=latex] (CA) -- (Q);
		\draw[<-, >=latex] (CA) -- (A);
		\draw[<-, >=latex] (A) -- (P);
	\end{tikzpicture}
\caption{Provenance} \label{fig:provgraph}
\end{figure}

The provenance of \k{e} includes both \k{parse} and \k{quote}. We pick an origin
by traversing the graph from the root, following the left-hand side member of
assignments until a leaf. It represents the provenance that is further modified
to yield the expression passed to \eval. In our example, the origin is \k{parse}.

% Wrap tables do not play well with lists (they warn about it in the documentation actually!)
\begin{wraptable}{r}{5cm}\small\centering
\begin{tabular}{lcc}\toprule
\bf Provenance & \bf \#sites & \bf \%sites \\\midrule
		\packageProvenanceNamea & \packageNbProvenanceSitesa & \packagePercentProvenanceSitesa \\
		\packageProvenanceNameb & \packageNbProvenanceSitesb & \packagePercentProvenanceSitesb \\
		\packageProvenanceNamec & \packageNbProvenanceSitesc & \packagePercentProvenanceSitesc \\
		\packageProvenanceNamee & \packageNbProvenanceSitese & \packagePercentProvenanceSitese \\
		\packageProvenanceNamef & \packageNbProvenanceSitesf & \packagePercentProvenanceSitesf \\
		\packageProvenanceNameg & \packageNbProvenanceSitesg & \packagePercentProvenanceSitesg \\
		\packageProvenanceNameh & \packageNbProvenanceSitesh & \packagePercentProvenanceSitesh \\
		\packageProvenanceNamei & \packageNbProvenanceSitesi & \packagePercentProvenanceSitesi \\
		\packageProvenanceNamel & \packageNbProvenanceSitesl & \packagePercentProvenanceSitesl \\
		\packageProvenanceNamen & \packageNbProvenanceSitesn & \packagePercentProvenanceSitesn \\\bottomrule
\end{tabular}
\caption{Origin} \label{tab:detailed_provenances}
\medskip
\begin{tabular}{rrr} \toprule
\bf Origin  & \bf \#sites & \bf \%sites \\\midrule
Reflection &  \packageNbReflectionSites & \packageReflectionSitesPercent\\
String & \packageNbStringSites & \packageStringSitesPercent \\
Constructed & \packageNbConstructedSites & \packageConstructedSitesPercent \\
Environment & \packageNbSymbolSites & \packageSymbolSitesPercent \\
External & \packageNbExternalSites & \packageExternalSitesPercent \\\bottomrule
\end{tabular}
\caption{Provenance}\label{tab:provenance}
\end{wraptable}

@@

\begin{itemize}[---]
\item {\it Reflection:} use of \k{match.call} to reflectively capture the
  expression that invoked this function, arguments are promises, and
  \k{substitute} is used to retrieve their source.
\item {\it String:} create from strings by invoking \k{parse},
  \k{str2expression}, or \k{str2lang}.
\item {\it Constructed:} invoked \k{quote} or \k{expression}, or by built with
  \k{call} or \k{as.call}.
\item {\it Environment: } created with \k{as.name} or \k{as.symbol}, typically
  to read a non-local environment.
\item {\it External: }  call to a C function. % Put an example
\end{itemize}

>>>>

Where does the expression passed to \eval come from? There are various means of
creating an expression, often these means correlate with a particular use case.
Our analysis records the values returned by some functions of interest as well
as their arguments. We build a direct acyclic graph that track the provenance
of the values that are given to \eval. Fig~\ref{fig:provgraph} shows an example
of the provenance graph for the expression \k{e}.
The provenance of \k{e} includes both \k{parse} and \k{quote}. We pick an
origin by traversing the graph from the root, following the left-hand side
member of assignments until a leaf. It represents the provenance that is
further modified to yield the expression passed to \eval. In our example, the
origin is \k{parse}.

\begin{figure}[H]
  \begin{minipage}{.39\textwidth}
    \vspace{-35mm}
\begin{lstlisting}
 > e <- parse(text = "a;b")
 > e[[2]] <- quote(c) # e is a;c
 > eval(e)
\end{lstlisting}
\end{minipage}
\begin{minipage}{.59\textwidth}
	\centering
	\begin{tikzpicture}[
		every node/.style={draw, rounded corners, fill=LightGray, text width = 2cm, align = center, thick}
		]
		\node (E) at (2,0) {\k{eval(e)}};
		\node[text width = 3cm] (CA) at (2, 1.5) {\k{e[[2]] <-quote(c)}};
		\node (Q) at (4, 3) {\k{quote(c)}};
		\node[text width = 4cm] (A) at (0, 3) {\k{e <- parse(text = "a;b")}};
		\node[text width = 3cm] (P) at (0, 4.5) {\k{parse(text = "a;b")}};

		\draw[<-, >=latex] (E) -- (CA);
		\draw[<-, >=latex] (CA) -- (Q);
		\draw[<-, >=latex] (CA) -- (A);
		\draw[<-, >=latex] (A) -- (P);
	\end{tikzpicture}
\end{minipage}
\caption{Example of a provenance graph} \label{fig:provgraph}
\end{figure}

% Wrap tables do not play well with lists (they warn about it in the documentation actually!)
\begin{wraptable}{r}{5cm}\small\centering
\begin{tabular}{lcc}\toprule
\bf Provenance & \bf \#sites & \bf \%sites \\\midrule
		\packageProvenanceNamea & \packageNbProvenanceSitesa & \packagePercentProvenanceSitesa \\
		\packageProvenanceNameb & \packageNbProvenanceSitesb & \packagePercentProvenanceSitesb \\
		\packageProvenanceNamec & \packageNbProvenanceSitesc & \packagePercentProvenanceSitesc \\
		\packageProvenanceNamee & \packageNbProvenanceSitese & \packagePercentProvenanceSitese \\
		\packageProvenanceNamef & \packageNbProvenanceSitesf & \packagePercentProvenanceSitesf \\
		\packageProvenanceNameg & \packageNbProvenanceSitesg & \packagePercentProvenanceSitesg \\
		\packageProvenanceNameh & \packageNbProvenanceSitesh & \packagePercentProvenanceSitesh \\
		\packageProvenanceNamei & \packageNbProvenanceSitesi & \packagePercentProvenanceSitesi \\
		\packageProvenanceNamel & \packageNbProvenanceSitesl & \packagePercentProvenanceSitesl \\
		\packageProvenanceNamen & \packageNbProvenanceSitesn & \packagePercentProvenanceSitesn \\\bottomrule
\end{tabular}
\caption{Origin} \label{tab:detailed_provenances}
\end{wraptable}

@@

\begin{itemize}[---]
\item {\it Reflection:} use of \k{match.call} to reflectively capture the
  expression that invoked this function, arguments are promises, and
  \k{substitute} is used to retrieve their source.
\item {\it String:} create from strings by invoking \k{parse},
  \k{str2lang}, or \k{str2expression}.
\item {\it Constructed:} invoked \k{quote} or \k{expression}, or by built with
  \k{call} or \k{as.call}.
\item {\it Environment: } created with \k{as.name} or \k{as.symbol}, typically
  to read a non-local environment.
\item {\it External: }  call to a C function. % Put an example
\end{itemize}

\begin{wraptable}{r}{5cm}\small\centering
\begin{tabular}{rrr} \toprule
\bf Origin  & \bf \#sites & \bf \%sites \\\midrule
Reflection &  \packageNbReflectionSites & \packageReflectionSitesPercent\\
String & \packageNbStringSites & \packageStringSitesPercent \\
Constructed & \packageNbConstructedSites & \packageConstructedSitesPercent \\
Environment & \packageNbSymbolSites & \packageSymbolSitesPercent \\
External & \packageNbExternalSites & \packageExternalSitesPercent \\\bottomrule
\end{tabular}
\caption{Provenance}\label{tab:provenance}
\end{wraptable}
